---
title: "DownSamp_NoiseRat"
output: html_document
---

```{r}
library(Seurat)
library(variancePartition)
library(edgeR)
library(scater)
library(ggplot2)
library(dplyr)
library(broom)
library(reshape2)
library(patchwork)
```


choose parameters (integration type, clustering res, min pct threshold)
```{r}
f<- 'Harmony.Batchindividual'
pct<-0.2
res<- 'SCT_snn_res.1'

```

```{r}
path<- here::here("output/mergedObjects/")

merged<- readRDS(paste0(path,f, ".rds"))
```

subset cells, run variance partition to see the effect of experiment size (total number of cells) on median variance explained by residuals

subsetting cells evenly between replicates and individuals

using clusters defined at res 1 (28 clusters), not subsetting equally between clusters

downsampling to total cell counts of 2700,5400,10800,21600 (only numbers divisible by 9, equal cell numbers from each group. only 2418 cells came from 19160 in Batch2, so the max cells we can test using this approach is ~21000 to still have equal cells from each individual/batch group)



```{r}
rep_subsamp<- function(ncells, nreps){
set.seed(1)



cellids<- list()


#take even numbers of cells from each replicate and individual (9 total groups)
npergroup<- ncells/9
metsub<- merged@meta.data

n<-0

repeat{
sizegroup<- c()
  for (k in 1:3){
    ind<- unique(metsub$individual)[k]
    sub<- metsub[metsub$individual == ind,]
    
    for(g in 1:3){
    repl<- unique(metsub$Batch)[g]
    cells<- rownames(sub[sub$Batch == repl,])
    
    samp<- sample(cells, npergroup, replace=FALSE)
    
   sizegroup<- c(sizegroup, samp)
   
  
  
    }
    
   
  }
 n<- n+1
    print(n)
    cellids[[n]]<- sizegroup

 if(n > (nreps-1)){
    break
    }
}

return(cellids)
}

```

```{r}
subone<- rep_subsamp(2700,10)
subtwo<- rep_subsamp(5400,10)
subthree<- rep_subsamp(7200,10)
subfour<- rep_subsamp(10800,10)
subfive<- rep_subsamp(16200,10)
subsix<- rep_subsamp(21600, 10)

```

```{r}
path<- here::here("output/GeneLists_by_minPCT/")
genelist<- read.table(file = paste0(path, "genelist.PCTthresh",pct,"_",f,".rds_",res,".txt"), sep=",")
genelist<- as.vector(genelist$x)

#subset merged to only the genes with PCT > min pct threshold in at least 1 cluster
mergesub<- subset(merged, features = genelist )
```



```{r}
median_exp_resids<- function(cellids){
varpart.list<- NULL
varpart.meds<- NULL
ncells.ind<- NULL
ncells.rep<- NULL
ncellsi<- NULL
ncellsr<- NULL

  for (i in 1:length(cellids)){
  print(i)
msub<- subset(mergesub, cells = cellids[[i]])

ncells.ind[[i]]<- table(msub@meta.data$SCT_snn_res.1, msub@meta.data$individual)
ncells.rep[[i]]<- table(msub@meta.data$SCT_snn_res.1, msub@meta.data$Batch)

ncellsi[[i]]<- ncells.ind
ncellsr[[i]]<- ncells.rep
#sub<- DGEList(counts=msub@assays$SCT@data, lib.size=colSums(msub@assays$SCT@data), samples=msub@meta.data)

  
  #group meta
  Group<- factor(paste(msub@meta.data[,res], msub@meta.data$Batch, msub@meta.data$individual, sep="."))

msub<- AddMetaData(msub, Group, col.name = "Group")

submerged<- as.SingleCellExperiment(msub, assay="SCT")
sumex<- sumCountsAcrossCells(submerged, ids=submerged@colData$Group)

Group<- colnames(sumex)
cluster<- as.vector(substr(Group, 1, regexpr("*.B", Group)-1))
batch<- substr(Group, regexpr("Batch", Group),regexpr("Batch", Group)+5)
ind<-  substr(Group, regexpr("NA", Group),regexpr("NA", Group)+6)

samps<- cbind(cluster,batch,ind,Group)

dge<- DGEList(sumex, samples=samps, remove.zeros = T)

 #remove ribosomal genes
  genes.ribo <- grep('^RP',rownames(dge),value=T)
  genes.no.ribo <- rownames(dge)[which(!(rownames(dge) %in%   genes.ribo))]
  dge$counts <- dge$counts[which(rownames(dge$counts) %in% genes.no.ribo),]
  
  
  #CalcNormFactors
  dge<- calcNormFactors(dge, method="TMM")
  
  #specify design matrix
  design<- model.matrix(~dge$samples$cluster+dge$samples$batch+dge$samples$ind)
  
  #voom
  v<- voom(dge, design, plot=F)
  #voom.plots[[i]]<- v
  
  #form
  form<- ~ (1|cluster) + (1|batch) + (1|ind)
  
  #run variance partition
  varpart<- suppressWarnings(fitExtractVarPartModel(v, form, dge$samples, useWeights=TRUE, quiet=TRUE, showWarnings = FALSE))
  
  #store varpart results
  varpart.meds[i]<- median(varpart$Residuals)
  
  
}

varpart.list<- list(varpart.meds, ncellsi, ncellsr)
return(varpart.list)

}
```

```{r}
remove(merged)
remove(genelist)
```

```{r,eval=F}
meds_subone<- median_exp_resids(subone)
```



```{r,eval=F}
saveRDS(meds_subone, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_2700cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
```


```{r,eval=F}
meds_subtwo<- median_exp_resids(subtwo)
saveRDS(meds_subtwo, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_5400cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
```

```{r,eval=F}
meds_subthree<- median_exp_resids(subthree)
saveRDS(meds_subthree, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_7200cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
```

```{r,eval=F}
meds_subfour<- median_exp_resids(subfour)
saveRDS(meds_subfour, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_10800cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
```

```{r,eval=F}
meds_subfive<- median_exp_resids(subfive)
saveRDS(meds_subfive, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_16200cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
```

```{r,eval=F}
meds_subsix<- median_exp_resids(subsix)
saveRDS(meds_subsix, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_21600cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
```

```{r}
meds_subone<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_2700cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
meds_subtwo<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_5400cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
meds_subthree<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_7200cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
meds_subfour<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_10800cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
meds_subfive<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_16200cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")
meds_subsix<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/downsamp_21600cells_10subreps_medianexplainedbyresiduals_varpart_PsB.rds")

```

```{r}
#get median of medians
objects<- list(meds_subone[[1]], meds_subtwo[[1]], meds_subthree[[1]], meds_subfour[[1]], meds_subfive[[1]], meds_subsix[[1]])
medmeds<- c()
for (i in 1:6){
  m<- median(objects[[i]])
  medmeds<- c(medmeds, m)
}

medmeds
```

boxplots of median variance explained by residuals in 10 subsets of cells at each experiment size
```{r}
s<- c(2700,5400,7200,10800,16200, 21600)
names(objects)<- s
boxplot(objects, xlab= "experiment size", ylab= "median variance explained by residuals")
```


```{r}
#reformat to dataframe
dat<- data.frame(s, medmeds)
colnames(dat)<- c("SampleSize", "MedianExp")

```

fit an exponential decay:
```{r}
decay_fit<- nls(MedianExp~SSasymp(SampleSize, yf, y0, log_alpha), data=dat)

decay_fit
```


```{r}
ggplot(dat, aes(x = SampleSize, y=MedianExp))+
  geom_point()+
  xlim(0,100000)+
  stat_smooth(method="nls", formula = y~ SSasymp(x, Asym, R0, lrc), se=FALSE, fullrange=TRUE)
```


```{r}
#median exp predicted for other total sample sizes (#s of high quality cells sequenced)

cellspersamp<- c(100,500,1000,2000,4000,10000)
cellstot<- cellspersamp*9

MedExpSizes<- NULL

for (i in 1:length(cellstot)){
  numb<- SSasymp(cellstot[i], 0.4112, 0.8112, -9.2141)
  MedExpSizes<- c(MedExpSizes, numb)
  
}

MedExpSizes
```

```{r}
dat<- data.frame(cellspersamp, cellstot, MedExpSizes)
```


```{r,eval=F}

write.csv(dat, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/ResidualVariances_fromDownSampAnalysis.csv")
```
  

visualize how many cells per cluster in each experiment size
```{r}
indbyclust<- list(meds_subone[[2]], meds_subtwo[[2]], meds_subthree[[2]], meds_subfour[[2]], meds_subfive[[2]], meds_subsix[[2]])
ncellsperclust<- NULL

for (i in 1:6){
  l<- indbyclust[[i]]
  clustcount<- NULL
  for (j in 1:10){
    su<- rowSums(l[[10]][[j]])
    clustcount<- cbind(clustcount, su)
  }
  ncellsperclust[[i]]<- clustcount
}
```

At each downsample size: plot ncells per cluster (yaxis: total cells and percent of total cells)


w/y axis total cells: 
```{r}
s<- c(2700,5400,7200,10800,16200, 21600)

totalcells.plots<- NULL
for (i in 1:6){
# prepare data
m<- melt(ncellsperclust[[i]])
m$Var1<- as.factor(m$Var1)

#plot
p<- ggplot(m, aes(x= Var1, y= value)) +
  geom_violin()+
  stat_summary(fun.y=median, geom="point", size=1,color="red")+#plot median as red dot
  ggtitle(paste0('ncells per cluster sampled from ', s[i], ' total cells')) +
  xlab("cluster")+
  ylab("number of cells")+
  scale_y_continuous(breaks=seq(0, s[i],s[i]/90))


totalcells.plots[[i]]<- p
}

totalcells.plots
```


power analysis using variance at each experiment size learned from downsampling analysis

adapted from code by Abhishek Sarkar and Anthony Hung.

```{r}
colors <- cbPalette <- c('#AA3377', "#E69F00", "#56B4E9", "#009E73", "grey50", "#0072B2", "#D55E00", "#CC79A7")
samp_size <- c(3, 10, 20, 40, 58, 100)
alpha <- 5e-6 #for FWER of 0.05

power_function <- function(x, n, v){ 
     pnorm(qnorm(alpha/2) + x * sqrt(n/v))
}

power.plots<- NULL

ns<- c(3,10, 30,58, 100)


  i<- 3
p <- ggplot(data.frame(x = c(0, 1)), aes(x = x)) +
     stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[1]),
                      aes(colour = "300"), size = .75)+
    stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[2]),
                      aes(colour = "1500"), size = .75)+
    stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[3]),
                      aes(colour = "3000"), size = .75)+
   stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[4]),
                      aes(colour = "6000"), size = .75)+
    stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[5]),
                      aes(colour = "12000"), size = .75)+
    stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[6]),
                      aes(colour = "30000"), size = .75)+
  scale_x_continuous(name = "Effect Size",
                              limits=c(0, 1)) +
     scale_y_continuous(name = "Power",
                           limits = c(0,1)) +
  theme_classic()+
     ggtitle(paste0("n = ", i)) +
     scale_colour_manual("Cells per individual", breaks = c("300", "1500", "3000", "6000", "12000", "30000"), values = colors) +
     #scale_linetype_manual("Curve Type", breaks = c("Power", "dynamic QTL FPR"), values = c("dynamic QTL FPR" = "dotted", "Power" = "solid")) +
     #theme_bw() +
     geom_hline(yintercept = .8, linetype = "dashed", color = "red")

power.plots[[1]]<- p

for (j in 2:5){
  i<- ns[j]
p <- ggplot(data.frame(x = c(0, 1)), aes(x = x)) +
     stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[1]),
                      aes(colour = "300"), size = .75)+
    stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[2]),
                      aes(colour = "1500"), size = .75)+
    stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[3]),
                      aes(colour = "3000"), size = .75)+
   stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[4]),
                      aes(colour = "6000"), size = .75)+
    stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[5]),
                      aes(colour = "12000"), size = .75)+
    stat_function(fun = power_function, args = list(n = i, v= dat$MedExpSizes[6]),
                      aes(colour = "30000"), size = .75)+
  scale_x_continuous(name = "Effect Size",
                              limits=c(0, 1)) +
     scale_y_continuous(name = element_blank(),limits = c(0,1)) +
  theme_classic()+
     ggtitle(paste0("n = ", i)) +
     scale_colour_manual("Cells per individual", breaks = c("300", "1500", "3000", "6000", "12000", "30000"), values = colors) +
     #scale_linetype_manual("Curve Type", breaks = c("Power", "dynamic QTL FPR"), values = c("dynamic QTL FPR" = "dotted", "Power" = "solid")) +
     #theme_bw() +
     geom_hline(yintercept = .8, linetype = "dashed", color = "red")

power.plots[[j]]<- p
}




power.plots
```

```{r}
V<- (power.plots[[1]]+ NoLegend())+(power.plots[[2]]+NoLegend())+(power.plots[[3]]+NoLegend())+(power.plots[[4]]+NoLegend())+power.plots[[5]] +plot_layout(ncol=5)
V
```

```{r, eval=F}
png(file= "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/figs/Fig5_DownSamp_Power.png", width=11, height=3, units= "in", res=1080)

V
  
dev.off()
```



if n=58, b=0.5, alpha=5e-6, 6000 cells per individual (collected across 3 batches) then power equals:

```{r}
power_function(0.6, 58, dat$MedExpSizes[3])
```

```{r}
sessionInfo()
```


