---
title: "Integrate lanes from all batches of EB pilot"
output: html_document
---


```{r, message=FALSE}
library(Seurat)
library(harmony)
library(ggplot2)
library(DataCombine)
library(here)
library(RColorBrewer)
options(future.globals.maxSize= 15000*1024^2) #allow global exceeding 4Gb
```


Read in the files, add metadata, and create an object list
```{r, warning=F}
filelist<-list.files(here::here('output/sampleQCrds/'), full.names = T)
objectlist<- list()
for (i in 1:length(filelist)){
  rds<- readRDS(filelist[i])
  objectlist[i]<- rds
  
}

```

create a merged seurat object
```{r}
ids<-substr(basename(filelist),1,12)
merged<- merge(objectlist[[1]], c(objectlist[[2]], objectlist[[3]],objectlist[[4]],objectlist[[5]],objectlist[[6]],objectlist[[7]],objectlist[[8]],objectlist[[9]],objectlist[[10]],objectlist[[11]],objectlist[[12]],objectlist[[13]],objectlist[[14]],objectlist[[15]],objectlist[[16]]),add.cell.ids=ids, merge.data=T)
```

```{r}
#need to fix the individual names because they are slightly different from batch1
replacements<- data.frame(from= c("SNG-NA18511.variant2", "SNG-NA18858.variant2", "SNG-NA19160.variant2"), to=c("SNG-NA18511", "SNG-NA18858", "SNG-NA19160"))
merged@meta.data<-FindReplace(merged@meta.data, "individual", replacements, from = "from", to= "to", exact=T, vector=F )
```

```{r}
#run PCA on full dataset pre-alignment
all.genes= rownames(merged)
merged<-FindVariableFeatures(merged,selection.method="vst", nfeatures = 5000)
#have previously used all genes (nfeatures=25000) and clustering by individual rather than batch (based on proportion of cells per cluster) was still observed downstream. Now using 5000 because it is the upper bound of what has been recommended in the literature.
merged<- ScaleData(merged, features = all.genes)
merged<-RunPCA(merged, npcs = 100, verbose=F)
```

```{r}
DimPlot(merged, reduction = "pca", group.by = "Batch")
```

Now, running harmony to integrate. Here, using Batch, SampleID(10x Lane), and individual to integrate. Since Batch and Lane are confounded, this may over correct for Batch.

```{r}
merged<- RunHarmony(merged, c("Batch", "individual"), plot_convergence = T, assay.use = "SCT")
```



Visualize Harmony embeddings
```{r}
DimPlot(merged, reduction="harmony", group.by= c("individual", "Batch"), combine=F)
```

Now Running UMAP and identifying clusters, etc

```{r}
merged<- RunUMAP(merged, reduction = "harmony", dims = 1:100, verbose = F)
merged<- FindNeighbors(merged, reduction="harmony", dims = 1:100, verbose = F)
merged<- FindClusters(merged, resolution=1, verbose = F)
merged<- FindClusters(merged, resolution=0.8, verbose = F)
merged<- FindClusters(merged, resolution=0.5, verbose = F)
merged<- FindClusters(merged, resolution=0.1, verbose = F)
```

SAVING merged/aligned/reclustered object
```{r, eval=F}
path<- here::here("output/mergedObjects/")
saveRDS(merged, file=paste0(path,'Harmony.Batchindividual.rds'))
```


```{r}
#reassign idents
Idents(merged)<- 'SCT_snn_res.1'
```

```{r}
VizDimLoadings(merged, dims = 1:2, reduction = "harmony")
VizDimLoadings(merged, dims = 3:4, reduction = "harmony")
VizDimLoadings(merged, dims = 5:6, reduction = "harmony")
```
```{r}
xlim <- c(min(merged@reductions$harmony@cell.embeddings[,'harmony_1']),
          max(merged@reductions$harmony@cell.embeddings[,'harmony_1']))
ylim <- c(min(merged@reductions$harmony@cell.embeddings[,'harmony_2']),
          max(merged@reductions$harmony@cell.embeddings[,'harmony_2']))

individuals <- table(merged$individual)
individuals <- individuals[individuals>50]
individuals <- names(individuals)
for (i in individuals)
{
  print(DimPlot(merged, reduction = "harmony", group.by = c("Batch"), pt.size = 0.01,
                cells = WhichCells(merged, expression = individual == i)) +
          xlim(xlim) + ylim(ylim) + ggtitle(i))
}
```



```{r}
DimPlot(merged, reduction = "umap")
DimPlot(merged, reduction = "umap", group.by = "Batch")
DimPlot(merged, reduction = "umap", group.by = "individual")
xlim <- c(min(merged@reductions$umap@cell.embeddings[,'UMAP_1']),
          max(merged@reductions$umap@cell.embeddings[,'UMAP_1']))
ylim <- c(min(merged@reductions$umap@cell.embeddings[,'UMAP_2']),
          max(merged@reductions$umap@cell.embeddings[,'UMAP_2']))
for (i in individuals)
{
  print(DimPlot(merged, reduction = "umap", 
                cells = WhichCells(merged, expression = individual == i)) +
          xlim(xlim) + ylim(ylim) + ggtitle(i))
}
```


```{r}
plots2<- DimPlot(merged, group.by = "individual", split.by = "Batch")
plots2
```
```{r}
DimPlot(merged, group.by = "Batch", split.by = c("individual"))
```




```{r}
DimPlot(merged, group.by = "SCT_snn_res.1", split.by = c("Batch"), label=T)
```

```{r}
DimPlot(merged, reduction = "harmony", group.by = "SCT_snn_res.1", split.by = "Batch", combine = F)
```


```{r}
VlnPlot(merged, features = c("POU5F1", "PAX6", "TNNT2", "SOX17", "HAND1", "LUM"), ncol=2)
```

```{r}
#pluripotent markers
FeaturePlot(merged, features = c("POU5F1", "SOX2", "NANOG"), pt.size = 0.2, ncol=3)
```

```{r}
#Endoderm markers (first 3 definitive endo, 4-6 liver markers, )
FeaturePlot(merged, features = c("SOX17","CLDN6","FOXA2", "TTR", "AFP", "FGB"), pt.size = 0.2, combine = F)


```

```{r}
#Mesoderm Markers (first 3 early meso markers, 4-6 heart markers, 7-9 endothelial markers (which comes from mesoderm), then some other general muscle markers)
FeaturePlot(merged, features = c("HAND1", "BMP4", "TNNT2","KDR", "GNG11", "ECSCR", "COL3A1", "ACTC1"), pt.size = 0.2, combine=F)
```

```{r}
#Ectoderm Markers (3-1 early ectoderm markers, 4-6schwann cell (myelinating, non myelinating, or precursor), 7-8 oligodendrocytes, 9-10 radial glia)
FeaturePlot(merged, features = c("PAX6", "GBX2",  "NES", "MPZ", "SOX10","GAP43", "OLIG1", "OLIG2", "VIM", "HES5"), pt.size = 0.2, ncol=3, combine=F)
```

```{r}
#More ectoderm, specifically neurons
#immature neurons: NEUROD1
#Mature Neurons: MAP2, SYP
#dopaminergic: TH, FOXA2,
FeaturePlot(merged, features = c("MAP2", "SYP","NEUROD1", "TH" ), pt.size = 0.2, ncol=3)

```

Identify cluster markers
```{r}
#how many cells per cluster?
t1<-table(merged@meta.data$SCT_snn_res.1, merged@meta.data$Batch)
t1
```


```{r}
#how many cells per cluster from each individual?
t2<-table(merged@meta.data$SCT_snn_res.1, merged@meta.data$individual)
t2
```


Reclustering with less resolution, check if everything is robust

```{r}
#reassign idents
Idents(merged)<- 'SCT_snn_res.0.5'
```

```{r}
DimPlot(merged, reduction = "umap")
DimPlot(merged, reduction = "umap", group.by = "Batch")
DimPlot(merged, reduction = "umap", group.by = "individual")
xlim <- c(min(merged@reductions$umap@cell.embeddings[,'UMAP_1']),
          max(merged@reductions$umap@cell.embeddings[,'UMAP_1']))
ylim <- c(min(merged@reductions$umap@cell.embeddings[,'UMAP_2']),
          max(merged@reductions$umap@cell.embeddings[,'UMAP_2']))
for (i in individuals)
{
  print(DimPlot(merged, reduction = "umap", 
                cells = WhichCells(merged, expression = individual == i)) +
          xlim(xlim) + ylim(ylim) + ggtitle(i))
}
```


```{r}
#reassign idents
Idents(merged)<- 'SCT_snn_res.0.1'
```

```{r}
DimPlot(merged, reduction = "umap")
DimPlot(merged, reduction = "umap", group.by = "Batch")
DimPlot(merged, reduction = "umap", group.by = "individual")
xlim <- c(min(merged@reductions$umap@cell.embeddings[,'UMAP_1']),
          max(merged@reductions$umap@cell.embeddings[,'UMAP_1']))
ylim <- c(min(merged@reductions$umap@cell.embeddings[,'UMAP_2']),
          max(merged@reductions$umap@cell.embeddings[,'UMAP_2']))
for (i in individuals)
{
  print(DimPlot(merged, reduction = "umap", 
                cells = WhichCells(merged, expression = individual == i)) +
          xlim(xlim) + ylim(ylim) + ggtitle(i))
}
```







```{r}
#reassign idents
Idents(merged)<- 'SCT_snn_res.0.8'
```

```{r}
DimPlot(merged, reduction = "umap")
DimPlot(merged, reduction = "umap", group.by = "Batch")
DimPlot(merged, reduction = "umap", group.by = "individual")
xlim <- c(min(merged@reductions$umap@cell.embeddings[,'UMAP_1']),
          max(merged@reductions$umap@cell.embeddings[,'UMAP_1']))
ylim <- c(min(merged@reductions$umap@cell.embeddings[,'UMAP_2']),
          max(merged@reductions$umap@cell.embeddings[,'UMAP_2']))
for (i in individuals)
{
  print(DimPlot(merged, reduction = "umap", 
                cells = WhichCells(merged, expression = individual == i)) +
          xlim(xlim) + ylim(ylim) + ggtitle(i))
}
```



```{r}
VlnPlot(merged, features= "percent.mt", group.by = "SCT_snn_res.1", pt.size = 0)
```

```{r}
merged[["percent.rps"]]<- PercentageFeatureSet(merged, pattern = "^RPS")
merged[["percent.rpl"]]<- PercentageFeatureSet(merged, pattern = "^RPL")
merged[["percent.rp"]]<- merged[["percent.rps"]]+merged[["percent.rpl"]]
VlnPlot(merged, features= "percent.rp", group.by = "SCT_snn_res.1", pt.size=0)
```

```{r}
FeaturePlot(merged, features = "nFeature_RNA")
```

```{r}
head(merged)
```


```{r}
VlnPlot(merged, features= "nFeature_RNA", group.by = "SCT_snn_res.1", pt.size=0)
```


```{r}
FeaturePlot(merged, features = c("POU5F1", "SOX17",  "HAND1", "PAX6"), pt.size = 0.2, ncol=2, combine=T)
```

```{r}
FeaturePlot(merged, features = c("FGB", "ECSCR",  "NEUROD1", "SOX10"), pt.size = 0.2, ncol=2)
```

```{r}
sessionInfo()
```

