---
title: "ReferenceIntegration"
output: html_document
---
```{r}
library(Seurat)
library(edgeR)
library(loomR)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(harmony)
library(ggplot2)
```

loading data

first, my data
```{r}
merged<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/mergedObjects/Harmony.Batchindividual.rds")
```



loading in hESC and iPS-to-EB raw dges from scHCL reference set

```{r}
hESC<- read.table("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/data/dge/hESC1.rawdge.txt.gz", header=T, row.names = 1)

iPStoEBday20<- read.table("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/data/dge/iPS-to-EB_20Day_dge.txt.gz", header=T, row.names = 1)
```
Note: there is no available metadata for these iPS to EB differentiations (no cell annotations online)



make seurat objects with all of the data
```{r}
hESC.obj<- CreateSeuratObject(hESC)
EB20.obj<- CreateSeuratObject(iPStoEBday20)
```

checking quality of cells
```{r}
FeatureScatter(hESC.obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
```

```{r}
#check MT%
#add %MT to metadata
hESC.obj[["percent.mt"]]<- PercentageFeatureSet(hESC.obj, pattern= "^MT-")
```

```{r}
FeatureScatter(hESC.obj, feature1 = "nCount_RNA", feature2 = "percent.mt")
```

```{r}
summary(hESC.obj$percent.mt)
```


```{r, eval=F}
#normalizing each scHCL dataset
hESC.obj<- suppressWarnings(SCTransform(hESC.obj, variable.features.n=5000,verbose=F))
EB20.obj<-suppressWarnings(SCTransform(EB20.obj, variable.features.n=5000,verbose=F))
```


load loom file from Cao et al (fetal tissue scRNA-seq)
```{r, eval=F}
Cao<- connect("/project2/gilad/katie/Pilot_HumanEBs/DataFromCaoetal202/GSE156793_S3_gene_count.loom", mode="r+")

Cao
```

```{r,eval=F}
Cao[["matrix"]]
```


```{r,eval=F}
#which genes are protein coding genes
protein<- which(Cao$row.attrs$gene_type[]=="protein_coding")

#sample genes of all main_cluster_name groups
#making a data fram of cell names and their cluster identity
cell.idents<- Cao$get.attribute.df(MARGIN = 2, col.names = "sample", attributes="Main_cluster_name")

cell.idents<- cell.idents %>% rownames_to_column("cellid")

#I'd like to sample ~500 cells from each cluster. Some clusters have fewer than 500 cells, so for those clusters I will take all cells. To do this, I make a dataframe listing the sample size of each cell type.
s.size<- table(cell.idents$Main_cluster_name)
s.size[s.size > 500]<- 500
#reorder
s.size<- s.size[unique(cell.idents$Main_cluster_name)]

#sample from each cluster
#make nested df
nest<- cell.idents %>% group_by(Main_cluster_name) %>% nest() %>% ungroup() %>% mutate(n = s.size)
 
sampled_nest<- nest %>% mutate(samp= map2(data, n, sample_n))

cells.keep<- sampled_nest %>% select(-data) %>% unnest(samp)
cells.keep<- cells.keep$cellid
cells<- which(cell.idents$cellid %in% cells.keep)
```

```{r,eval=F}
#get metrix for cells and genes of interest
Cao.matrix<- Cao[["matrix"]][cells,protein]
```

```{r,eval=F}
#get metadata
attrs<- c("Assay", "Batch", "Experiment_batch", "Main_cluster_name", "Fetus_id", "Sex")
attr.df<-Cao$get.attribute.df(MARGIN = 2, col.names = "sample", attributes=attrs)

#subset the attr.df to just the subset of cells
attr.df<- attr.df[cells,]
```

```{r,eval=F}
#name matrix rows and columns
rownames(Cao.matrix)<- rownames(attr.df)
colnames(Cao.matrix)<- Cao[["row_attrs/gene_short_name"]][protein]

#seems like there are some non-unique gene names, which is expected but annoying
dups<- which(duplicated(colnames(Cao.matrix)))

length(dups)
```


```{r,eval=F}
#sum the counts for duplicate gene names
Cao.matrix.test<- (t(Cao.matrix))
rowname<- rownames(Cao.matrix.test)
Cao.matrix.test<- cbind(rowname, as.data.frame(Cao.matrix.test))

rownames(Cao.matrix.test)<- NULL


Cao.agg.dups<- aggregate(.~rowname, Cao.matrix.test, sum)

anyDuplicated(Cao.agg.dups$rowname)
```

```{r,eval=F}
#rename rows
rownames(Cao.agg.dups)<- Cao.agg.dups$rowname

#remove rowname column
Cao.agg.dups<- Cao.agg.dups[, 2:ncol(Cao.agg.dups)]

colnames(Cao.agg.dups)<- rownames(attr.df)
```

Now, Convert to Seurat Object. Major question: is this raw data? or are these counts normalized?
```{r,eval=F}
Cao.obj<- CreateSeuratObject(Cao.agg.dups, meta.data= attr.df)
```


Normalize the fetal reference set
```{r,eval=F}
#first normalizing with SCTransform
Cao.obj<- SCTransform(Cao.obj, variable.features.n=5000,verbose=F)
```

```{r, eval=F}
saveRDS(Cao.obj,"/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/CaoEtAl.Obj.CellsOfAllClusters.ProteinCodingGenes.rds" )
```

```{r}
Cao.obj<-readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/CaoEtAl.Obj.CellsOfAllClusters.ProteinCodingGenes.rds")
```

```{r}

#check features, QC between batches/individuals
#Cao data has no MT genes
FeatureScatter(Cao.obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "Fetus_id")
```
```{r}
FeatureScatter(Cao.obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "Batch")
```

```{r}
FeatureScatter(Cao.obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "Experiment_batch")
```
```{r}
summary(Cao.obj@meta.data$nCount_RNA)
```
```{r}
summary(Cao.obj@meta.data$nFeature_RNA)
```

```{r}
Cao.obj[["percent.mt"]]<- PercentageFeatureSet(Cao.obj, pattern= "^MT-")
```

```{r}
summary(Cao.obj@meta.data$percent.mt)
```

```{r}
Cao.obj<- RunPCA(Cao.obj, npcs= 100, verbose = F)
```

```{r}
DimPlot(Cao.obj, reduction = "pca")
DimPlot(Cao.obj, reduction = "pca", group.by = "Fetus_id")
DimPlot(Cao.obj, reduction = "pca", group.by = "Experiment_batch")
DimPlot(Cao.obj, reduction = "pca", group.by = "Batch")
DimPlot(Cao.obj, reduction = "pca", group.by = "Main_cluster_name")
VizDimLoadings(Cao.obj, dims = 1:2, reduction = "pca")
VizDimLoadings(Cao.obj, dims = 3:4, reduction = "pca")
VizDimLoadings(Cao.obj, dims = 5:6, reduction = "pca")
```
```{r}
Cao.obj<- FindNeighbors(Cao.obj, dims = 1:30, verbose = F)
Cao.obj<- RunUMAP(Cao.obj, dims=1:30, verbose = F)

DimPlot(Cao.obj, group.by = "Main_cluster_name", label = T, label.size = 2) + NoLegend()
```
```{r}
DimPlot(Cao.obj, group.by = "Fetus_id")
```

```{r}
DimPlot(Cao.obj, group.by = "Experiment_batch")
```

```{r}
#rename Cao metadata so none match with EB object (just need to replace Batch)
colnames(Cao.obj@meta.data)

colnames(Cao.obj@meta.data)[5]<-"Batch_week"
```

```{r, eval=F}
#rename orig.idents
hESC.obj$orig.ident<- "scHCL.hESC"
EB20.obj$orig.ident<- "scHCL.EB20"
merged$orig.ident<- "EB.Pilot"
Cao.obj$orig.ident<- "Cao.EtAl"
```

```{r, eval=F}
#merge objects
obj.list<- list(Cao.obj, merged, hESC.obj, EB20.obj)
merge.all<- merge(x=obj.list[[1]], y=c(obj.list[[2]], obj.list[[3]], obj.list[[4]]), merge.data=T)
```

```{r, eval=F}
merge.all<- SCTransform(merge.all, variable.features.n = 5000, vars.to.regress = c("orig.ident"), assay= "SCT")
```

```{r, eval=F}
all.genes= rownames(merge.all)
merge.all<-FindVariableFeatures(merge.all,selection.method="vst", nfeatures = 5000)
merge.all<- ScaleData(merge.all, features = all.genes, assay = "SCT")
merge.all<-RunPCA(merge.all, npcs = 100, verbose=F, Assay="SCT")
```


```{r, eval=F}
merge.all<- RunHarmony(merge.all, c("orig.ident", "individual", "Batch"), theta = c(3,1,1), plot_convergence = T, assay.use = "SCT")
```

```{r, eval=F}
merge.all<- RunUMAP(merge.all,dims=1:100, reduction="harmony")
```

```{r, eval=F}
saveRDS(merge.all,"/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/merge.all.SCTwRegressOrigIdent.Harmony.rds" )
```

```{r}
merge.all<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/merge.all.SCTwRegressOrigIdent.Harmony.rds")
```

```{r}
V<- DimPlot(merge.all, group.by = "Main_cluster_name", label = T, label.size = 2.5, repel = T)+NoLegend()

V
```
```{r, eval=F}
pdf(file = "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/pdfs/IntegrateReference_UMAP_MainClusterID_SCTregress.pdf")

V

dev.off()
```

Add new metadata to include Main_cluster_name as well as cluster labels from EB Pilot DE results, and orig ident of the HCL data
```{r}
merge.all<- AddMetaData(merge.all, col.name = "all.labels", metadata = merge.all@meta.data$Main_cluster_name)

#for cells with NA for main_cluster_name, replace label from SCT_snn_res 0.1
for (i in (1:nrow(merge.all@meta.data))){
if (is.na(merge.all@meta.data$all.labels[i]) == T){
  merge.all@meta.data$all.labels[i]<- merge.all@meta.data$SCT_snn_res.0.1[i]
}
}

#and now replace remaining NAs with orig.ident
for (i in (1:nrow(merge.all@meta.data))){
if (is.na(merge.all@meta.data$all.labels[i]) == T){
  merge.all@meta.data$all.labels[i]<- merge.all@meta.data$orig.ident[i]
}
}
```

```{r}
V<- DimPlot(merge.all, group.by = "all.labels", label = T, label.size = 2.5, repel = T)+NoLegend()

V
```

```{r, eval=F}
pdf(file = "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/pdfs/IntegrateReference_UMAP_allLabels_SCTregress.pdf")

V

dev.off()
```

```{r}
V<- DimPlot(merge.all, group.by = "orig.ident", label = F, order= c("scHCL.hESC","scHCL.EB20", "EB.Pilot","Cao.EtAl"))
  

V
```

```{r, eval=F}
pdf(file = "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/pdfs/IntegrateReference_UMAP_origident_all_SCTregress.pdf")

V

dev.off()
```

```{r}
DimPlot(merge.all, split.by = "orig.ident", label = F, order= c("scHCL.hESC","scHCL.EB20", "EB.Pilot","Cao.EtAl"))
```

```{r}
options(ggrepel.max.overlaps = Inf)
```

```{r}
#subset object to just my data and Cao reference, plot UMAP
Idents(merge.all)<- "orig.ident"
sub<- subset(merge.all, idents= c("Cao.EtAl", "EB.Pilot"))
V<- DimPlot(sub, group.by = "all.labels", label = T, repel = T, label.size = 2.5)+NoLegend()
V
```

```{r, eval=F}
pdf(file = "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/pdfs/IntegrateReference_UMAP_EBpilotAndCao_SCTregress.pdf")

V

dev.off()
```

```{r}
V<- DimPlot(sub, group.by = "Main_cluster_name", label = T, repel = T, label.size = 2.5)+NoLegend()
V
```

```{r, eval=F}
pdf(file = "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/pdfs/IntegrateReference_UMAP_EBpilotAndCao_groupbyMainClusterName_SCTregress.pdf")

V

dev.off()
```

```{r}
V<- DimPlot(sub, group.by = "SCT_snn_res.0.1", label = T, repel = T, label.size = 2.5)+NoLegend()
V
```


```{r, eval=F}
pdf(file = "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/pdfs/IntegrateReference_UMAP_EBpilotAndCao_groupbySeuratClusterRes0.1_SCTregress.pdf")

V

dev.off()
```

```{r}
DimPlot(sub, group.by = "orig.ident")
```

```{r}
DimPlot(sub, group.by = "orig.ident", order="Cao.EtAl")
```

```{r}
#Subset object to just my data and HCL references, plot UMAP
sub<- subset(merge.all, idents= c("scHCL.EB20", "EB.Pilot"))
V<- DimPlot(sub, group.by = "orig.ident", label = F)
V
```


```{r}
sub<- subset(merge.all, idents= c("scHCL.EB20", "EB.Pilot", "Cao.EtAl"))
DimPlot(sub, group.by="orig.ident", pt.size = 0.2, label=F) 
```

```{r}
FeaturePlot(merge.all, split.by = "orig.ident", features = c("TERF1", "POU5F1", "DPPA3"), slot = "data", pt.size = 0.1)
```

```{r}
DimPlot(merge.all, split.by="orig.ident",group.by = "all.labels", pt.size = 0.2, label=F) +NoLegend()
```



Now, will transfer labels for Cao Et Al and hESC onto my data. 
```{r, eval=F}
#subset to remove scHCL.EB20
sub<- subset(merge.all, idents= c("Cao.EtAl", "EB.Pilot", "scHCL.hESC"))
#compute distance matrix based on harmony embeddings, dims 1:100
har_embeds<- sub@reductions$harmony@cell.embeddings

har_distmat<- as.matrix(dist(har_embeds, method="euclidean", upper=TRUE))
```


```{r, eval=F}
#vectors with cell ids from Cao.EtAl, EB.pilot, and scHCL.hESC
EB.pilot.ids<-rownames(merge.all@meta.data[merge.all@meta.data$orig.ident == "EB.Pilot",])

#subset rows to only cells in EB.pilot
sub_har_distmat<- har_distmat[rownames(har_distmat) %in% EB.pilot.ids,]

#subset cols to only cells not in EB.pilot
'%notin%'<- Negate('%in%')
sub_har_distmat<- sub_har_distmat[,colnames(sub_har_distmat) %notin% EB.pilot.ids]

```

```{r, eval=F}
nearest.ref.cell.id<- NULL
nearest.ref.cell.dist<- NULL
#for loop, loop through each row
for (i in 1:nrow(sub_har_distmat)){
  nearest.ref.cell.dist[i]<- min(sub_har_distmat[i,])
  nearest.ref.cell.id[i]<- names(which.min(sub_har_distmat[i,]))
}

nearest.ref.table<- cbind(rownames(sub_har_distmat), nearest.ref.cell.id,nearest.ref.cell.dist)

colnames(nearest.ref.table)<- c("EB.cell.id", "nearest.ref.cell.id", "harmony.dist.to.nearest.ref.cell")
```

```{r, eval=F}
#add annotation
ann<- as.data.frame(merge.all@meta.data$all.labels)
ann<- cbind(rownames(merge.all@meta.data), ann)
colnames(ann)<- c("nearest.ref.cell.id", "annotation")

nearest.ref.table<- as.data.frame(nearest.ref.table)
nearest.ann<- left_join(nearest.ref.table, ann, by=c("nearest.ref.cell.id"))

write.csv(nearest.ann, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/NearestReferenceCell.Cao.hESC.EuclideanDistanceinHarmonySpace.csv")
```

```{r}
nearest.ann<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/NearestReferenceCell.Cao.hESC.EuclideanDistanceinHarmonySpace.csv", header=T, row.names = 1)
```


```{r, eval=F}
a<- as.data.frame(table(nearest.ann$annotation))
a<- a[a$Var1 != "scHCL.EB20",]
a<- a[a$Var1 != "0",]
a<- a[a$Var1 != "1",]
a<- a[a$Var1 != "2",]
a<- a[a$Var1 != "3",]
a<- a[a$Var1 != "4",]
a<- a[a$Var1 != "5",]
a<- a[a$Var1 != "6",]
a<- a[a$Var1 != "7",]
colnames(a)<- c("reference.cell.type", "Frequency")

write.csv(a, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/NearestReferenceCell.Cao.hESC.FrequencyofEachAnnotation.csv")
```

```{r}
a<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/NearestReferenceCell.Cao.hESC.FrequencyofEachAnnotation.csv", header=T, row.names = 1)
```


```{r}
sub<- subset(merge.all, idents= c("EB.Pilot"))
EB.cell.id<- rownames(sub@meta.data)
sub@meta.data<- cbind(sub@meta.data, EB.cell.id)
sub@meta.data<- full_join(sub@meta.data, nearest.ann, by= c("EB.cell.id"))
rownames(sub@meta.data)<- EB.cell.id

V<- DimPlot(sub, group.by="annotation", pt.size = 0.2, label.size = 2.5,label=T, repel=T) +NoLegend()

V
```
```{r, eval=F}
pdf(file = "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/pdfs/IntegrateReference_UMAP_EBpilot_annotationsFromReference.pdf")

V

dev.off()
```

Instead of matching to just one nearest cell, we get a more robust annotation if we check a group of nearest neighbors for common annotations
```{r, eval=F}
mostcommon.ann<- NULL
maxann.FIVEnearest<- NULL

#for loop, loop through each row
for (i in 1:nrow(sub_har_distmat)){
  cell<- sub_har_distmat[i,]
  cell<- cell[order(cell)]
  topfive<- names(cell[1:5])
  
  #get the annotations of the nearest 5 reference cells
  topfiveann<- merge.all@meta.data$all.labels[rownames(merge.all@meta.data) %in% topfive]
  
  #if/else at least 3/5 match annotations
  maxann<- max(table(topfiveann))
  finalann<- names(which.max(table(topfiveann)))
  
  maxann.FIVEnearest[i]<- maxann
  
  if(maxann >= 3){
    mostcommon.ann[i]<- finalann
  } else {
    mostcommon.ann[i]<- "uncertain"
  }
  
}

CommonAnnDF<- as.data.frame(cbind(rownames(sub_har_distmat), mostcommon.ann, maxann.FIVEnearest))
colnames(CommonAnnDF)<- c("EB.cell.id", "Annotation", "NoutofFIVErefneighborsWithSameAnnotation")
```

```{r, eval=F}
write.csv(CommonAnnDF, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/MostCommonAnnotation.FiveNearestRefCells.csv")
```

```{r}
CommonAnnDF<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/MostCommonAnnotation.FiveNearestRefCells.csv", header=T, row.names = 1)
```


```{r, eval=F}
b<- table(CommonAnnDF$Annotation)

write.csv(b, "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/Frequency.MostCommonAnnotation.FiveNearestRefCells.csv")
```

```{r}
b<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/Frequency.MostCommonAnnotation.FiveNearestRefCells.csv", header=T, row.names = 1)
```


```{r}
#print fetal celltypes not present in EB data
'%notin%'<- Negate('%in%')
unique(merge.all@meta.data$Main_cluster_name)[unique(merge.all@meta.data$Main_cluster_name) %notin% b$Var1] 
```


```{r}
sub<- subset(merge.all, idents= c("EB.Pilot"))
EB.cell.id<- rownames(sub@meta.data)
sub@meta.data<- cbind(sub@meta.data, EB.cell.id)
sub@meta.data<- full_join(sub@meta.data, CommonAnnDF, by= c("EB.cell.id"))
rownames(sub@meta.data)<- EB.cell.id

V<- DimPlot(sub, group.by="Annotation", pt.size = 0.2, label.size = 2.5,label=T, repel=T) +NoLegend()

V
```

```{r, eval=F}
pdf(file = "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/pdfs/IntegrateReference_UMAP_EBpilot_CommonAnnotationsFromFiveReferenceNeighbors.pdf")

V

dev.off()
```

 

```{r}
sessionInfo()
```


