---
title: "ReferenceAnn_DE"
output: html_document
---

```{r}
library(Seurat)
library(limma)
library(edgeR)
library(Matrix)
library(dplyr)
library(scater)
```

```{r}
merged<- readRDS("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/mergedObjects/Harmony.Batchindividual.rds")
```

```{r}
#load annotations
ann<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/MostCommonAnnotation.FiveNearestRefCells.csv")
```

```{r}
merged@meta.data$RefAnn<- ann$Annotation
```

```{r}
DimPlot(merged, group.by= "RefAnn")+NoLegend()
```

```{r}
clusts<- (unique(merged@meta.data$RefAnn))

#theres 1 thymocyte and it doesnt want to plot it. so removing
clusts<- clusts[-57]

#or myeloid
clusts<- clusts[-58]

#or excitatory neurons
clusts<- clusts[-66]

#or antigen presenting
clusts<- clusts[-65]

plts<- NULL
for (i in 1:length(clusts)){
  b<- clusts[i]
  plts[[i]]<- DimPlot(merged, cells= c(rownames(merged@meta.data[merged@meta.data$RefAnn == as.character(b),])))+NoLegend()+ggtitle(paste0(b))
}

plts
```


```{r}
#Add Group Metadata
Group<- factor(paste(merged@meta.data$RefAnn, merged@meta.data$Batch, merged@meta.data$individual, sep="."))

merged<- AddMetaData(merged, Group, col.name = "Group")
```


subset to:
cells from groups that have at least 10 cells total
```{r}
Ngroup<- table(merged@meta.data$Group)
Ngroup.keep<- Ngroup[Ngroup>=10]

merged<- subset(merged, subset= Group %in% names(Ngroup.keep))
```


```{r}
submerged<- as.SingleCellExperiment(merged, assay="RNA")
```

```{r}
sumex<- sumCountsAcrossCells(submerged, ids=submerged@colData$Group)

sumex<- sumex[,colSums(sumex)>0]
```


```{r}
Group<- colnames(sumex)
cluster<- as.vector(substr(Group, 1, regexpr("*.B", Group)-1))
batch<- substr(Group, regexpr("Batch", Group),regexpr("Batch", Group)+5)
ind<-  substr(Group, regexpr("NA", Group),regexpr("NA", Group)+6)

samps<- cbind(cluster,batch,ind,Group)
```

```{r}
dge<- DGEList(sumex, samples=samps, remove.zeros = T)
```

remove uncertain
```{r}
dge<-dge[,dge$samples$cluster != "uncertain"]
```

keep cell types that have at least 4 groups (at least 2 individuals, at least 2 replicates)
```{r}
ct.keep<- table(dge$samples$cluster)
ct.keep<- names(ct.keep[ct.keep>=4])

dge<- dge[,dge$samples$cluster %in% ct.keep]
```

23 cell types left after this cell/celltype filtering



```{r}
#filtering genes

genes.keep<- filterByExpr(dge, group= Group)

dge<- dge[genes.keep,,keep.lib.sizes=F]

```

```{r}
dge<- calcNormFactors(dge, method="TMM")
```

```{r}
dge$samples$cluster<- as.factor(as.character(dge$samples$cluster))
```

```{r}
#not including batch and individual in model because I do not have balances representation across cell types

design<- model.matrix(~0+ dge$samples$cluster)
```

```{r}
v<- voom(dge, design, plot=T)
```




```{r}
fit<- lmFit(v,design)
```


```{r}
nclust<- length(unique(dge$samples$cluster))
contrasts<- NULL
for (i in 1:nclust){
    c<- c(rep(-1,nclust))
    c[i]<- nclust-1
    
    contrasts<- cbind(contrasts, c)
}
```

```{r}
contrasts
```
```{r}
fit<- contrasts.fit(fit, contrasts=contrasts)
```

```{r}
efit<- eBayes(fit)
```

```{r}
plotSA(efit)
```

```{r}
summary(decideTests(efit))
```

```{r}
levels(dge$samples$cluster)
```

For example, lets see the top genes DE by Cardiomyocytes (coef 2)
```{r}
topTable(efit, coef=2, sort.by="P", n=50)
```
We see recognizeable markers of CMs in that list-- MYL7, MYL4, TNNT2

```{r}
vol<- topTable(efit, coef=2, n=nrow(fit))
labsig<- vol[rownames(vol) %in% c("MYL7", "MYL4", "TNNT2"),]
labsiggenes<- rownames(labsig)
thresh<- vol$adj.P.Val < 0.05
vol<-cbind(vol, thresh)
v<- ggplot(vol, aes(x=logFC, y= -log10(adj.P.Val))) +
  geom_point(aes(colour=thresh), show.legend = FALSE) +
  scale_colour_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  geom_text(data=labsig, aes(label=labsiggenes))

v
```

```{r, eval=F}
png(file= "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/figs/Supp_RefAnnDE_cardiomyocyte.png", width= 6, height=6, units= "in", res= 1080)

v

dev.off()
```

make list of top 20 most significant DE genes in each of the 23 cell types in this analysis
```{r}
Top20DE<- NULL

for (i in 1:nclust){
  Top20DE[[i]]<- topTable(efit, coef=i, sort.by="P", n=50)
}

names(Top20DE)<- levels(dge$samples$cluster)

Top20DE
```



checking for top DE markers from Cao and descartes browser

```{r}
#metanephric: EYA1, UNCX, LRP2, NKAIN3, TMEM132D
 tab<- topTable(efit, coef=14, sort.by="P", n=2000)
tab[rownames(tab) %in% c("EYA1", "UNCX", "LRP2", "NKAIN3", "TMEM132D", "IGF2", "AUTS2", "DACH1", "ROBO2", "H19", "PTPRG", "ERBB4", "NCAM", "FZD7", "SIX2", "OSR1", "WT1", "PAX2", "MEOX1", "MEIS1", "MEIS2"),]
```
```{r}
FeaturePlot(merged, features=c("EYA1", "UNCX", "LRP2", "NKAIN3", "TMEM132D"))
```

```{r}
#hepatoblasts: SLC2A25, ACSS2, ASSI. SLC2A25 and ASS1 are not in our data set
 tab<- topTable(efit, coef=8, sort.by="P", n=2000)
tab[rownames(tab) %in% c("SLC2A25", "ACSS2", "ASSI", "DLK1", "KRT18", "KRT8"),]
```

```{r}
vol<- topTable(efit, coef=8, n=nrow(fit))
labsig<- vol[rownames(vol) %in% c("AFP", "FGB", "ACSS2"),]
labsiggenes<- rownames(labsig)
thresh<- vol$adj.P.Val < 0.05
vol<-cbind(vol, thresh)
b<- ggplot(vol, aes(x=logFC, y= -log10(adj.P.Val))) +
  geom_point(aes(colour=thresh), show.legend = FALSE) +
  scale_colour_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  geom_text(data=labsig, aes(label=labsiggenes))

b
```

```{r, eval=F}
png(file= "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/figs/Supp_RefAnnDE_hepatoblast.png", width= 6, height=6, units= "in", res= 1080)

b

dev.off()
```

```{r}
#amacrine: MEIS2
 tab<- topTable(efit, coef=1, sort.by="P", n=2000)
tab[rownames(tab) %in% c("MEIS2"),]
```




mesothelial
```{r}
vol<- topTable(efit, coef=13, n=nrow(fit))
labsig<- vol[rownames(vol) %in% c("NID2", "COL6A3", "COL1A1", "COL3A1", "COL6A1"),]
labsiggenes<- rownames(labsig)
thresh<- vol$adj.P.Val < 0.05
vol<-cbind(vol, thresh)
p<- ggplot(vol, aes(x=logFC, y= -log10(adj.P.Val))) +
  geom_point(aes(colour=thresh), show.legend = FALSE) +
  scale_colour_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  geom_text(data=labsig, aes(label=labsiggenes))

p
```

```{r, eval=F}
png(file= "/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/figs/Supp_RefAnnDE_mesothelial.png", width= 6, height=6, units= "in", res= 1080)

p

dev.off()
```


```{r}
sessionInfo()
```

