---
title: "RefSubIntegrationComparison"
output: html_document
---
```{r}
library(dplyr)
```



```{r}
plur<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/TranferredAnnotations_ReferenceInt_JustPluripotent.csv")
end<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/TranferredAnnotations_ReferenceInt_JustEndoderm.csv")
mes<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/TranferredAnnotations_ReferenceInt_JustMeso.csv")
ee<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/TranferredAnnotations_ReferenceInt_JustEarlyEcto.csv")
neur<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/TranferredAnnotations_ReferenceInt_JustNeuron.csv")
nc<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/TranferredAnnotations_ReferenceInt_JustNeuralCrest.csv")
```

```{r}
sepint<- rbind(plur, end, mes, ee, neur, nc)
colnames(sepint)<- c("X.sep", "EB.cell.id", "Annotation.sep", "NofFive.sep")
```


```{r}
all<- read.csv("/project2/gilad/katie/Pilot_HumanEBs/Embryoid_Body_Pilot_Workflowr/output/MostCommonAnnotation.FiveNearestRefCells.csv")
```

```{r}
comb<- full_join(all, sepint, by = "EB.cell.id")

dim(comb)
```

```{r}
sum(as.character(comb$Annotation) == as.character(comb$Annotation.sep))
```

```{r}
33714/42488
```

79% of cells are assigned to the same cell type between full integration and partial integration


Let's see if we see any patterns in the cell types that disagree between integrations
```{r}
dis<- comb[as.character(comb$Annotation) != as.character(comb$Annotation.sep),]
```

```{r}
barplot(table(dis$Annotation), las=2, cex.names = 0.4)
```

```{r}
barplot(table(dis$Annotation.sep), las=2, cex.names = 0.4)
```

Lots of cells flipping to and from hESC, uncertain. 


How many unique cell types total in the separate integrations?

```{r}
length(unique(sepint$Annotation.sep))
```

how many in full integration?
```{r}
length(unique(all$Annotation))
```

Both include uncertain. so its 66 fetal cell types + hESC in total, 60 fetal cell types + hESC from sep int. 

Which cell types do we lose?

```{r}
septypes<- as.character(unique(sepint$Annotation.sep))
alltypes<- as.character(unique(all$Annotation))

'%notin%' <- Negate('%in%')
lost<- alltypes[(alltypes %notin% septypes)]

lost
```
How many of these cell types were there in the full integration? were they rare?

```{r}
taball<- table(all$Annotation)

taball[names(taball) %in% lost]
```

looks like there are some cell types in the separate integration that are not in the full



```{r}
gain<- septypes[septypes %notin% alltypes]
gain
```

```{r}
tabsep<- table(sepint$Annotation.sep)

tabsep[names(tabsep) %in% gain]
```

see most common label transitions

```{r}
trans<- paste(dis$Annotation, "_", dis$Annotation.sep)

length(unique(trans))
```

```{r}
switch<- sort(table(trans), decreasing = T)
switch
```

how many are switching to or from hESC or uncertain?

```{r}
length(grep("hESC|uncertain", trans))
```

```{r}
7190/length(trans)
```

Take a closer look at the other mismatches
```{r}
mis<- trans[-grep("hESC|uncertain", trans)]

mistab<- sort(table(mis), decreasing=T)

mistab
```

```{r}
sessionInfo()
```

